SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design

### 2025-04-07 src/main.tex
[text] double blind-ing
[text] compression

\title{%
  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
  Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
}

@@

\author{Quentin Guilloteau}
\orcid{0009-0003-7645-5044}
\email{Quentin.Guilloteau@inria.fr}
\affiliation{%
  \institution{Inria, CNRS, ENSL, UCBL1, LIP}
  \city{Lyon}
  \country{France}
}

\author{Antoine Waehren}
\email{A.Waehren@stud.unibas.ch}
\affiliation{%
  \institution{University of Basel}
  \city{Basel}
  \country{Switzerland}
}

\author{Florina M. Ciorba}
\email{Florina.Ciorba@unibas.ch}
\affiliation{%
  \institution{University of Basel}
  \city{Basel}
  \country{Switzerland}
}

@@

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months, from October 1$^{st}$ 2024 to April 1$^{st}$ 2025.
% \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}

@@

For example, authors can link to a Zenodo archive that will \emph{always} point to the latest version of an artifact.
Hence, the content of the artifact behind this link can change over time.

@@

\paragraph{Details}

\begin{itemize}
\item We ensure that an empty local Docker cache is used before building the containers.
\item Each container is allocated a 10-minute time frame to build.
\item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
\end{itemize}

@@

We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the Euro-Par 2024 conference that used a Docker container to share their software environments, and used \ecg\ to build their associated \df s once every month for seven months.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12" "2025-04-01"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
The deadline for the authors to submit their artifacts to Euro-Par 2024 was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
Hence, our initial build took place almost 5 months after the artifacts submission.

The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
The data collected and presented in this section is available on Zenodo~\cite{dataset_europar24}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{./figs/results_per_artifact.pdf}
  \caption{
    Evolution of the packages in the software environment of each container through time.
    Each container has been rebuilt once a month.
    The color of the bar corresponds to the month when a specific version of a package has been introduced in the software environment.
    We can see that the proportion of package versions similar to the versions in the initial build is decreasing through the months.
  }
  \label{fig:results_per_artifact}
\end{figure*}

>>>>

\title{%
  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
}

@@

% \author{Quentin Guilloteau}
% \orcid{0009-0003-7645-5044}
% \email{Quentin.Guilloteau@inria.fr}
% \affiliation{%
%   \institution{Inria, CNRS, ENSL, UCBL1, LIP}
%   \city{Lyon}
%   \country{France}
% }
% 
% \author{Antoine Waehren}
% \email{A.Waehren@stud.unibas.ch}
% \affiliation{%
%   \institution{University of Basel}
%   \city{Basel}
%   \country{Switzerland}
% }
% 
% \author{Florina M. Ciorba}
% \email{Florina.Ciorba@unibas.ch}
% \affiliation{%
%   \institution{University of Basel}
%   \city{Basel}
%   \country{Switzerland}
% }

\author{Anonymous Authors}
\email{Anonymous email}
\affiliation{%
  \institution{Anonymous Institution}
  \city{Anonymous City}
  \country{Anonymous Country}
}

@@

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
In practice, about 12\% of the artifacts use such tools~\cite{guilloteau2024longevity}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months, from October 1$^{st}$ 2024 to April 1$^{st}$ 2025.
We focus only on Docker containers as they are the most used in artifacts~\cite{guilloteau2024longevity}.
% \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}

@@

For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.% of an artifact.
Hence, the content of the artifact behind this link can change over time.

@@

\paragraph{Details}
\begin{itemize}
\item We ensure that an empty local Docker cache is used before building the containers.
\item Each container is allocated a 10-minute time frame to build.
\item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
\end{itemize}

@@

We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the Euro-Par 2024 conference that used a Docker container to share their software environments, and used \ecg\ to build their associated \df s once every month for seven months.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12" "2025-04-01"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
The deadline for the authors to submit their artifacts to Euro-Par 2024 was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
Hence, our initial build took place almost 5 months after the artifacts submission.
The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
The data collected and presented in this section is available on Zenodo~\cite{dataset_europar24}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{./figs/results_per_artifact.pdf}
  \caption{
    Evolution of the packages in the software environment of each container over time.
    %Each container has been rebuilt once a month.
    The color of the bar corresponds to the month when a specific version of a package has been introduced in the software environment.
    We can see that the decrease of the proportion of package versions similar to the versions in the initial build over time.
  }
  \label{fig:results_per_artifact}
\end{figure*}
